---
title: "Learning Apply Family Functions in R"
author: |
  | Neeraj Jain
  | neeraj2409@outlook.com
date: "24/03/2020"
output:
  html_document: default
  html_notebook: default
---
The objective of this worksheet is to enhance knowledge of the `apply` family functions. 

## 1. `apply` function
`apply` : apply function to each column or row of matrix (more accurately `array`). Since, dataframe is also matrix, therefore, `apply` also works on dataframe. See example,

```{r}
Df <- data.frame(a = 1:10, b = rnorm(10), c = runif(10, 10, 100))
apply(Df, 1, sum) # one as second argument indicate provide sum of each row.
apply(Df, 2, sum) # provide sum of each column

#More example: 
#Summary statistics of each column
apply(Df, 2, summary) 

#only mean of each column
apply(Df, 2, mean)

#min of each column
apply(Df, 2, min)

#sum of square of each column 
apply(Df, 2, function(x) sum(x^2))

#Number of NA in each column
apply(Df, 2, function(x) sum(is.na(x)))

```

**Try following excercise:**
```{r, echo=TRUE, error = TRUE}
# Ex1: Why sum of first 2 column is not returned? 
#Create following R object: 
Dfm <- data.frame(a = 1:10, b = 10:1, letters[1:10])
apply(Dfm, 2, sum)

#Ex2: Get mean and variance using single line function for each column. 
Df <- data.frame(a = 1:10, b = rnorm(10), c = runif(10, 10, 100))

#Ex3: Replace NA's value with 0 in following dataframe using single line function using apply
Df <- data.frame(a = c(1:9, NA), b = sample(c(1, 5, NA), size = 10, replace = TRUE))
````

```{r eval=FALSE, include=FALSE}
#without apply
Df[is.na(Df)] <- 0
apply(Df, 2, function(x){x[is.na(x)] <- 0
return(x)})
```

## 2. `lapply` function
`lapply` apply a Function to each element of a List or Vector. Lapply returns output in the list, of the same length of X (input vector or list). 

```{r}
v <- c(1, 2, 3, 4, 5)

#square root of each element
lapply(v, sqrt)

#square of each element
lapply(v, function(x) x^2)

#Function on list
Ls <- list(a = 1:10, b = runif(5), c = rnorm(15))

#length of each element
lapply(Ls, length)

#mean of each element
lapply(Ls, mean)

#summary statistics of each element of  list
lapply(Ls, summary)
```

**Try following excercise:**
```{r}
#Ex1: Why following code works on dataframe? 
Df <- data.frame(a = 1:10, b = 10:1, c = rnorm(10)) 
lapply(Df, mean)

#Ex2: calculate number of missing values for each column of data in list (Lsd). 
data(cars)
data("mtcars")
Lsd <- list(cars = cars, mtcars = mtcars)

#Ex3: Calculate summary of each column of data in Lsd. 

#Ex4: Write a function that use lapply within a lapply. 

#Ex5: Attempt all `apply` excercise using lapply
````

```{r, include=FALSE}
lapply(Lsd, FUN = function(x) apply(x, 2, function(y) sum(is.na(y))))
```

## 3. `sapply` function
Sapply is same as `lapply`, but it simplifies the output to vector or matrix, if possible. `sapply(x, f, simplify = FALSE, USE.NAMES = FALSE)` is the same as `lapply(x, f)`. Same example using `sapply`, and see the difference. 

```{r}
v <- c(1, 2, 3, 4, 5)

#square root of each element
sapply(v, sqrt)

#square of each element
sapply(v, function(x) x^2)

#Function on list
Ls <- list(a = 1:10, b = runif(5), c = rnorm(15))

#length of each element
sapply(Ls, length)

#mean of each element
sapply(Ls, mean)

#summary statistics of each element of  list
sapply(Ls, summary) #may be much better. 
```

Attempt all excercise of lapply using sapply and check the differences in the outcome. If you find yourself typing `unlist(lapply(...))`, stop and consider `sapply`.

## 4. `vapply` function
vapply is similar to sapply, but has a pre-specified `type` of return value, so it can be safer (and sometimes faster) to use. For example, 

```{r, error = TRUE}
vapply(Df, FUN = mean, FUN.VALUE = 0) #since return values are numeric, it works. 

vapply(Df, FUN = mean, FUN.VALUE = "a")  #does not work
```
For more information, please refer `?vapply`. 

## 5. `mapply` function
This is useful when you have several data structures (e.g. vectors, lists) and you want to apply a function to the 1st elements of each, and then the 2nd elements of each, etc., coercing the result to a vector/array as in `sapply`.

```{r, eval=TRUE, include=TRUE}
#Sums the 1st elements, the 2nd elements, etc. 
mapply(sum, 1:5, 1:5, 1:5) 

mapply(rep, 1:4, 4:1)   

#To generate random numbers with different mean and standard deviation
mapply(FUN = function(x, y) rnorm(5, x, y), 1:5, 5:1)
```
*I never felt the need of `mapply` function for any of the problem*.




