---
title: "Subsetting"
author: "Neeraj Jain"
date: "27-01-2020"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

**Subsetting**: can be used to return subsets of a vector, martix, or data frame which meet a particular condition. There are three methods of subsetting in R: 

###  1. **By Index**: `[ ]` (closed brackets) are used to subset elements from vector, matrix, daatframe or list.
  
### 1.1 **Atomic Vector**
```{r, error = TRUE}
rm(list = ls())

a <- c(1,2,3,4)

a[1] #getting first element
a[c(1,2)] #to get more than one elements
a[5] #returns NA as fifth element is not known. 
a[c(1, 1, 1)] #first element 3 times. 
a[c(1, 2, NA, 4)] #NA returns NA

#Negative index removes the elements. 
a[-2] #remove 2nd element
a[c(-3, -4)]  #to remove more than one elements. 


#another Example
a <- c(23, 24, 31, 31)
a <- a[-2] #remove 2nd element and overewrite a
a
 
#Assignment 
a[2] <- 12 #replacing 2 element from 12
a[c(1,2)] <- c(34, 101) #replacing first 2 element of with new values
a

#character Vector
BB <- c("Vedansh", "Asim", "Sehnaz", "Paras", "Abhijit", "Shukla")

BB[-c(1, 5)] #removing 1st and 5th element from BB

#adding more elements. Like adding "Aarti", and "Maraya" to BB
BB[c(5, 6)] <- c("aarti", "marya") #adding two names to BB
( BB <- c(BB, "aarti", "marya") )#alternative of last line #it is better too. 
length(BB)

BB[5] <- "Tarang" #replacing fifth element from "Tarang"

( BB <- c(BB[1:5], "Katrina Kaif", BB[6:10]) )#Adding another name between the character. 

#Another Example
SG <- c("Kukku", "Bunty", "Sartaz", "Gaitonde")

( SG <- c(SG[1:2], "Gurteg", SG[3:4]) ) #Adding another element between the SG
```


###  1.2 **Matrix** : Since, matrix has 2 dimension, thefore we used `,` to provide index to second argument. See example below: 
```{r, error = TRUE}
( Mat2 <- matrix(1:16, ncol = 4) )

Mat2[5] #fifth element of matrix. (raed column wise)

#element of 2nd row and 3rd column
Mat2[2,3] #, menas providing index for 2nd dimension (ie column)

#Be careful. `,` is used to provide index for second dimension of matrix. However, for atomic vetocr
a <- 1:4
a[1,2] #it wonot work. Since, a has only one dimension. 

#check the output of followings: 
Mat2[c(2,3) , ]
Mat2[ , c(2,3)]
Mat2[c(2,3), c(3,4)]
```

### To add column and rows in matrix use these two function: `rbind` an `cbind`

```{r, error = TRUE}
Mat2
rbind(Mat2, c(1,2,3,4))
cbind(Mat2, c(1,2,3,4))
```

### 1.3 **list**
```{r, error = TRUE}
( ls <- list(a = 1:4, b = BB, c = Mat2) )

ls[c(2, 3)] #to get 2nd and 3rd elements of list

ls2 <- ls[c(2)]
str(ls2) #output in list

ls4 <- ls[[2]] #double closed bracked to get element of list
str(ls4)
length(ls4)

#Always use single index in double closed bracket. 
ls[[2, 3]] #unpredictable resutls. Mostly returns error. 

#alternative of [[ is $
ls$b
ls$c

#we can do further subsetting on subsetting results. Like: 
ls$c[3, c(1,2)] 
ls$c[3, ][c(1,2)] #subsetting on subsetting results. 
```

### 1.4 **Dataframe**
```{r}
set.seed(123)
Df <- data.frame(a = rnorm(10), b = rnorm(10))

#like matrix (when use ,)
Df[1,2]
Df[1, ]
Df[, 1]
Df[c(7,8), 1]

#like list
Df$a[c(7,8)] 


str(Df[1])
str(Df[[1]])
```

### Exercise
```{r}
#Create following Df
set.seed(007)
Df2 <- as.data.frame(matrix(rnorm(16), ncol = 4))
```
 1.  Add new row as c(1,2,3, 4)
 2. First get 2nd column using subsetting, then calculate sum, var of this.
 3. Get 2nd column of your dataframe using 3 differetn methods. 
 4. Get sum of 3rd row. 
 5. Replace first column with 1:5



***********
### 2. **Subsetting by Logical value**
```{r}
rm(list = ls())

#logical Operators in R
a <- c(2, 4, 6, 10) 

a > 3
a < 3
a >= 3
a <= 3
a == 3
a != 3
```

### There are two logical function in R, 1) `all`, it is alternative of excel `and`, and 2) `any`, it is alternative of excel `or`. See example below

```{r}
all(TRUE,TRUE, TRUE, FALSE)
any(TRUE, FALSE, FALSE)

b <- c("a", "a ", "a")
b[c(1,2,3)] == "a"
all(b == "a") #to check all elements are equals to "a"

d <- c(rep(1, 100), 2, 2, rep(1, 200))
all(d == 1)
```

### Another useful function in R is `which`. It returns index of all `TRUE` value. A simple example is: `which(T, F, T)`, will retrun `#1 3`. 

```{r}
which(c(TRUE, FALSE, FALSE, TRUE, TRUE)) 

d[which(d != 1)] #getting values from index 

e <- c(TRUE, FALSE,TRUE)
!e  #! reverse the T and F

which(!(d == 1))  #do not use !d == 1

g <- c(2, 5, 7, 10, 6)
g[c(TRUE, FALSE, TRUE, FALSE, TRUE)]

#all values greater than 5
g > 5
g[g > 5]

#Small Example:
#a. values <= 2
#b.  values not equal to 7
#c.  how many's values are greater than 3

g[g <= 2] #a. 
g[g != 7] #b. 
sum(g > 3) #c. 
```
### There are two more logical opertors, 1) `&` and 2) `|`. `&`. These are like mathematical operators and works in the same way. For example, `TRUE & TRUE` returns `TRUE`, ie both elements on right and left are true, then `&` returns true. `|` is called or. If any of the element either on left or right is right, then `|` returns true. `|` returns `FALSE` only if both the elements are `FALSE` on right and left.  See more example: 

```{r}
c(T, F, T, F) | c(F, T, T, F)  #| is or
c(T, F, T, F) & c(F, T, T, F)
```

### Class Exercise
```{r}
#Question
set.seed(123) 
marks <- rnorm(82, mean = 65, sd = 20)
```

1. Validate your data to ensure that al values are between 0-100. If any value greter than 100, replace by 100, and in case of less than 0, replace by 0.

2. Percentage of students who have scored more than 60 marks. 
3. Get index of students who scored more than 90 marks, and also vector of marks.  
4. Get mean of top 10 and bottom 10 students marks. [Hint: order]
5. Percentage of stuents who have scored marks between 75 - 90. 
6. If gender of the students is given as follows: 
```{r}
Gender <- sample(x = c("F", "G"), size = 82, replace = TRUE)
```

7. Find number of male students, and female students. 
8. Find mean, sd, min, and max of males and female marks respectively.


************

3. **By elements name**
Check Lab 5.

