---
title: "Loop and Control flow"
author: "Neeraj Jain"
date: "17-02-2020"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
##1. Loop and control flow function (if)
```{r, include=FALSE}
rm(list =ls())
```

###1. Loop
```{r}
#Creating simple Loop in R
#Example: 1 - Print
for(i in 1:9) {
  print(i^2)
} 
#For each i, entire body of the loop will be evaluated in a sequence. 


#Example: 2 - Mean of each column without apply
mat <- matrix(1:16, ncol = 4)

apply(mat, 2, mean) #with apply

#Manually (one by one)
mean(mat[, 1])
mean(mat[, 2])
mean(mat[, 3])
mean(mat[, 4])

#using loop
a <- c()
for(i in 1:4) {
    a[i] <- mean(mat[, i]) 
}

#Example: 3 - aggregate using Loop
#Generating some data
set.seed(123)
n <- c("Rishu", "Rajat", "Udit", "Dhawal", "Manan")
Df <- data.frame(Names = sample(x = n, size = 30, replace = TRUE), 
                 Marks = rnorm(n = 30, mean = 10, sd = 3))
head(Df)
str(Df)

#Requirement: using loop, get mean marks and max marks of each name in `n`.
#Manually (for one student)
mean(Df[Df$Names == "Dhawal", 2])
max(Df[Df$Names == "Dhawal", 2])

# max(Df[Df$Names == "Dhawal", ]$Marks) #another way of subsetting

#using loop
Mean <- c() #creating empty mean vector
Max <- c() #creating empty max vector
for(i in 1:length(n)) {
  D <- Df[Df$Names == n[i], ]$Marks
  Mean[i] <- mean(D)
  Max[i] <- max(D)
}

(Summary <- data.frame(Names  = n, Mean1 = Mean, Max1 = Max))
```

##Loop Exercise: Plot using Loop
```{r}
#generating Data
set.seed(123)
Matrix <- matrix(data = rnorm(100), ncol =10)
colnames(Matrix) <- LETTERS[1:10]
```
Assume first column is dependend variable, and all other variables are independent variable. Do:

1. Plot each indenpendent variable against dependent variable. 

2. Save all plot in specific directory. Try to use names such that it convey information about the variable. 

3. Title in the plot should be "Scatterplot: A ~ `name of IV`"


##2. Control Flow: `if` function

The function has major two part. The first is condition `if(condition)`, and second is body (usually written in curly bracket). The expression in the body of `if` function is evaluated only if the condition in `if` is `TRUE`. If condition is not true, then the expression in `if` body is not evaluated at all. Take very simple example: 
```{r, error = TRUE}
if(TRUE) {
  a <- 2
}
a

if(FALSE) {
  b <- 2
}
b
```

Another example is square root. 
```{r}
#Square root (if number is positive)
srt <- function(x) {
  if(x  > 0 ) {
    x^0.5
  } else {
  print("Number is negative")
  }
}

#example: 
srt(10)
srt(-2)

#Annuity Function
Annuity <- function(A = 1, i, n) {
  if(i == 0) {
    n*A
  } else {
  A * (1 - (1+i)^(-n) ) / i
  }
}
#example
Annuity(i = .10, n = 10)
Annuity(i = 0 , n = 10)

#Annuity Function for different payments
Annuity <- function(A = 1, i, n, type = c("Monthly","HY", "Q", "A")) {
  if(n < 0) {
    stop("N is negative")
  }
  Fac <- c(Monthly = 12, HY = 2, Q = 4, A = 1)
  s <- Fac[type] 
  if(i == 0) {
    n*A*s
  } else {
    A * (1 - (1+i/s)^(-n*s) ) / (i/s)
  }
}

#Example
Annuity(10000, i = 0, n = 10, type = "Q")
Annuity(100, i = .10, n = 5, type = "Monthly")
```

